Chatbots and messaging platforms in the classroom: an analysis from the teacher's perspective

### 2021-12-02 edubots_surveys.Rnw
Responds to question on support

for the first time in their classes. Aggregating the answers, currently 91\% of
the teachers are using messaging apps as learning tools.

@@

Regarding the social factor of chat groups that teachers use in class, according
to Fig. \ref{figure:chatbots_messagingorganization} the vast majority prefer
small groups only with the students from the same course. These are more focused
groups with specific goals and dedicated to the organization of the course and
its tasks, and from the pedagogic point of view \textcolor{blue}{it also seems more appropriate to improve
the learning process}. Interestingly, and as a cross-check of the answers above,
about 30\% of teachers consider they should not be part of the chat group. This
might seem to contradict the results of the other survey, but in fact, being a
result of different surveys, to a certain point affirms the same thing: there is
a great amount of teachers that would be against being in a chat group with
students. Whether that's a slim majority or not, that's open to debate (and
would probably need a more pointed survey focused on this), but what is clear is
that forcing teachers to create chat groups with students and participate in
them would create a certain amount of resistance. Also, only university teachers
find interesting a group with all the students and teachers in their own Faculty
or School. The lack of teachers from vocational education here may be the
consequence of using specific language such as ``Faculty''. Moreover, the fact
that many universities and schools already use these groups for administrative
and social interaction (e.g. \cite{dibitonto2018chatbot}) might be the reason
for the low percentage of teachers that selected this response.

@@

From a qualitative point of view, these results show consensus in a few key aspects:
\begin{itemize}
 \item Teachers want to have support from their corresponding institution. This
   is probably due to the strict European data regulations and to avoid the
   overhead required to \textcolor{blue}{sign up new students every year}.
 \item Teachers need sustainability. The main reason is that implementing
   changes is expensive in terms of effort and prone to errors in the first
   years of implementing new procedures, therefore, it is logical that if
   chatbots or messaging platforms are to be introduced, they want this change
   to be as permanent as possible.
 \item Keep it Simple. Media interaction requirements are simple and only a 2.4\% is requesting more features. This could be a sign of technological burn out produced by the previous and current courses where, due to the pandemic, the use of computer has increased significantly.
\end{itemize}

An analysis of survey questions related to RQ4 seems to indicate that teachers
prefer to have interaction with their students in a chat "room"; a preference of
{\em horizontal} interaction might indicate the need to delegate or offload part
of the burden of answering, and attending, every single question posed by
students. In an horizontal setting, other students can answer if the teacher
does not do it immediately. This might also be the reason why the possibility of
using tools from which analytics can be extracted is also valued by teachers:
that way, activity in the chat tool, answering questions, regularity, can all be
observed and valued as part of the process of assessing achievement of learning
objectives by students.

>>>>

for the first time in their classes. % Aggregating the answers, currently 91\% of
% the teachers are using messaging apps as learning tools.

@@

Regarding the social factor of chat groups that teachers use in class, according
to Fig. \ref{figure:chatbots_messagingorganization} the vast majority prefer
small groups only with the students from the same course. These are more focused
groups with specific goals and dedicated to the organization of the course and
its tasks, and from the pedagogic point of view \textcolor{blue}{it also seems more appropriate to improve
the learning process}. Interestingly, and as a cross-check of the answers above,
about 30\% of teachers consider they should not be part of the chat group. This
might seem to contradict the results of the other survey, but in fact, being a
result of different surveys, to a certain point affirms the same thing: there is
a great amount of teachers that would be against being in a chat group with
students. \textcolor{blue}{However, students will still need to get the services
that the university, through chatbots, provide, thus opening the door for implantation of chatbots
without the intervention of the professor, simply tapping university provided
services \cite{bernier2002services}.}
Whether that's a slim majority or not, that's open to debate (and
would probably need a more pointed survey focused on this), but what is clear is
that forcing teachers to create chat groups with students and participate in
them would create a certain amount of resistance. Also, only university teachers
find interesting a group with all the students and teachers in their own Faculty
or School. The lack of teachers from vocational education here may be the
consequence of using specific language such as ``Faculty''. Moreover, the fact
that many universities and schools already use these groups for administrative
and social interaction (e.g. \cite{dibitonto2018chatbot}) might be the reason
for the low percentage of teachers that selected this response.

@@

From a qualitative point of view, these results show consensus in a few key aspects:
\begin{itemize}
 \item Teachers want to have support from their corresponding institution. This
   is probably due to the strict European data regulations \textcolor{blue}{on data protection},
   gut also to avoid the overhead required to \textcolor{blue}{sign up new students every year.
   In general support from specific formation to access to IT help desk for specific features
   enhances the possibility of adoption by the teachers or the possibility of a successful
   adoption that improves the learning outcomes. This is supported by the result of the two
   surveys: they prefer whatever "messaging" application is provided by the university, and
   also they value sustainability as well as "officiality" of the platform, as shown in
   Table \ref{tab:interaction}}.
 \item Teachers need sustainability. The main reason is that implementing
   changes is expensive in terms of effort and prone to errors in the first
   years of implementing new procedures, therefore, it is logical that if
   chatbots or messaging platforms are to be introduced, they want this change
   to be as permanent as possible.
 \item Keep it Simple. Media interaction requirements are simple and only a 2.4\% is requesting more features. This could be a sign of technological burn out produced by the previous and current courses where, due to the pandemic, the use of computer has increased significantly.
\end{itemize}

An analysis of survey questions related to RQ4 seems to indicate that teachers
prefer to \textcolor{blue}{share an interaction space with their student, as 
as shown in Tables \ref{tab:kindinteraction} and \ref{tab:interaction}, where a majority of responders
indicate they {\em don't} want unidirectional communication};a preference of
{\em horizontal} interaction, \textcolor{blue}{but also the answer to the kind of
chatbots preferred that features FAQ bots prominently, which might indicate a} 
need to delegate or offload part
of the burden of answering, and attending, every single question posed by
students. In an horizontal setting, other students can answer if the teacher
does not do it immediately. This might also be the reason why the possibility of
using tools from which analytics can be extracted is also valued by teachers:
that way, activity in the chat tool, answering questions, regularity, can all be
observed and valued as part of the process of assessing achievement of learning
objectives by students.
